---
title: "Assignment 8"
author: "Scott Shepard"
date: "6/1/2019"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prompt

This assignment is dependent on completion of previous assignment. 
Objective of this project is to forecast CME seat prices using CME trading
volume.

Now that you have a monthly time series of CME seat prices, the file
`Contracts_Volume.csv` contains the monthly trading volume for 
different CME products.  

Each row has the first date of the following month (the trade volume is 
for the month), the commodity indicator for CME, the short description of 
the commodity, whether it is Future (F) or Option (O), 
the electronic trading volume and the total trading volume. 
The floor trading volume is the difference between the total 
volume and the electronic volume.

The file `Contracts_Classification.csv` associates each CME commodity 
with a seat class.

Note that owning a CME seat gives a trader advantage for floor trading.
In contrast, owning a CME seat does not have any advantage for electronic
trading. Since electronic trading constitutes market participants trading
directly with other participants via CME's network connections and order
matching engine.

There are three classes of seats CME, IMM, and IOM which confer the 
rights to trade different sets of commodities traded. 
CME seat owners can trade everything, IMM seat owners are allowed to 
trade all futures and options except agricultural products, 
and IOM seat owners are allowed to trade future and options 
only in index products and all other CME options.

To find the volume, for example index products, add both the 
futures and options volume for the same index product for a month.

 
Your task is to use the trading volume information to forecast the CME
monthly seat price for 2013.  It is recommended to do exploratory data
analysis to find initial data relationships such as correlations. For
example, the total trade volume for all CME products might be a good
predictor for CME seat class, but not for the others. 
You may have to choose and select which commodities have influence on 
the IMM and IOM seat prices.

The tasks are outlined below.

Task A:

Use the following algorithms:

- Linear regression (seat price is independent, volume(s) dependent)
- Linear regression with ARMA errors (use arima with xreg)
- Holt Winters
- ARIMA
- Seasonal ARIMA (SARIMA) - here seasonality is monthly
- Fractional ARIMA (ARFIMA) - check applicability first using the ACF
- ARMA and GARCH combination - use the fGarch R library and garchFit()

Note that you have to implement each of the above algorithms for each of the 3 classes of seats: CME, IMM, IOM. 

Task B:

Since you already have the seat prices for the 2013,
evaluate each algorithm from Task A using sMAPE. 
Which one do you recommend to forecast monthly prices for each of the seat classes?

# Data

## Create Time Serieses

This function is copied over from assignment 7.

```{r, warning=F, message=F}
cmeS <- read.csv("~/Datasets/31006/cmeS.csv")
immS <- read.csv("~/Datasets/31006/immS.csv")
iomS <- read.csv("~/Datasets/31006/iomS.csv")

library(dplyr)
library(zoo)
library(ggplot2)
library(reshape2)

create_timeseries <- function(df_, interpolation='spline') {
  # Date formatting
  df_$DateOfSale <- as.Date(df_$DateOfSale, '%m/%d/%Y')
  df_$yearmon <- as.yearmon(df_$DateOfSale)
  
  # Group by month
  df_ <- df_ %>%
    group_by(yearmon) %>%
    summarize(price=mean(price)) %>%
    data.frame() %>%
    merge(data.frame(yearmon = as.yearmon(2001 + seq(0, 12*13-1)/12)),all=T)
  
  # Interpolate
  df_$interpolated <- is.na(df_$price)
  x = df_$yearmon
  if('constant' %in% interpolation) {
    df_$constant <- approx(df_$yearmon, df_$price, xout=x, method = "constant")$y  
  }
  if('linear' %in% interpolation) {
    df_$linear <- approx(df_$yearmon, df_$price, xout=x)$y
  }
  if('spline' %in% interpolation) {
    df_$spline <- spline(df_, n=nrow(df_))$y
  }
  df_
}

cmeS <- create_timeseries(cmeS)
immS <- create_timeseries(immS)
iomS <- create_timeseries(iomS)
```

## Contract Volume Data

The new datasets need to be merged in with the time series.
  1. Contracts Volume
  2. Contracts Classification

```{r, message=FALSE, echo=FALSE}
library(stringr)

cv <- read.csv("~/Datasets/31006/Contracts_Volume.csv", stringsAsFactors = F)
cc <- read.csv("~/Datasets/31006/Contracts_Classification.csv", stringsAsFactors = F)

cv$Electronic.Volume <- str_replace_all(cv$Electronic.Volume, ',', '')
cv$Electronic.Volume <- as.numeric(cv$Electronic.Volume)
cv$Floor.Volume <- cv$Total.Volume - cv$Electronic.Volume

df = merge(cv, cc, by.x='Commodity.Indicator', by.y='Commodity.Code', all.x=T)

df$Date <- as.Date(df$Date, '%m/%d/%Y')
df$yearmon <- as.yearmon(df$Date)

# IMM and IOM can only trade the commodities listed under those seats
imm_iom_monthly_vol <- df %>% 
  dplyr::filter(Division %in% c('IMM', 'IOM')) %>%
  group_by(yearmon, Division) %>%
  summarize(electric.vol = sum(Electronic.Volume), 
            total.vol = sum(Total.Volume), 
            floor.vol=sum(Floor.Volume))

# CME can trade all commodities, not just the ones listed
cme_monthly_vol <- df %>%
  group_by(yearmon) %>%
  summarize(electric.vol = sum(Electronic.Volume), 
            total.vol = sum(Total.Volume), 
            floor.vol=sum(Floor.Volume)) %>%
  mutate(Division = 'CME')

monthly_vol <- rbind(data.frame(cme_monthly_vol), data.frame(imm_iom_monthly_vol))

ggplot(monthly_vol, aes(x=yearmon, y=floor.vol, color=Division)) + 
  geom_line() + 
  labs(title='Monthly Floor Volume at the Merch',
       y='Floor Volume', x='Time')
```

```{r, warning=F, message=F}
attach_monthly_vol <- function(ts_df, division, monthly_df = monthly_vol) {
  monthly_df <- dplyr::filter(monthly_df, Division==division)
  monthly_df <- select(monthly_df, -Division)
  left_join(ts_df, monthly_df, by='yearmon')
}

ts_cmeS_all <- attach_monthly_vol(cmeS, 'CME')
ts_immS_all <- attach_monthly_vol(immS, 'IMM')
ts_iomS_all <- attach_monthly_vol(iomS, 'IOM')

head(ts_cmeS_all)
```

### Train / Test

Before fitting any models, remove the 2013 year from the timeseries objects
and keep the 2013 year as the values to judge the forecasts.

```{r, warning=F, message=F}
ts_cmeS_2013 <- dplyr::filter(ts_cmeS_all, yearmon >= 2013)
ts_immS_2013 <- dplyr::filter(ts_immS_all, yearmon >= 2013)
ts_iomS_2013 <- dplyr::filter(ts_iomS_all, yearmon >= 2013)

ts_cmeS <- dplyr::filter(ts_cmeS_all, yearmon < 2013)
ts_immS <- dplyr::filter(ts_immS_all, yearmon < 2013)
ts_iomS <- dplyr::filter(ts_iomS_all, yearmon < 2013)
```

# Task A: Fit Models

## Linear Regression 

Seat price is independent, Volume dependent

### CME

```{r, warning=F, message=F}
lm_cme <- lm(spline ~ floor.vol, data=ts_cmeS)
summary(lm_cme)

ts_floor_vol_scatter <- function(ts_df, division) {
  ggplot(ts_df, aes(x=floor.vol, y=spline)) + 
    geom_point() + 
    geom_smooth(method='lm') + 
    labs(title=paste(division, 'Seat Floor Volume vs. Interpolated Price'),
         subitle='Linear Model')
}

ts_with_floor_vol_plot <- function(ts_df, division) {
  X <- ts_df$yearmon
  Y1 <- ts_df$spline
  Y2 <- ts_df$floor.vol
  par(mar = c(5, 5, 3, 5))
  plot(x=X, y=Y1, type ="l", 
       ylab = paste("Interpolated Price of",division,"Seat"),
       main = paste(division, "Seat price & Floor Volume"), 
       xlab = "Time", xaxt='n',
       col = "blue")
  par(new = TRUE)
  plot(x=X, y=Y2, type = "l", 
       xaxt = "n", yaxt = "n", ylab = "", xlab = "", 
       col = "red", lty = 2)
  axis(side = 4)
  mtext("Floor Volume", side = 4, line = 3)
  legend("topleft", c("Seat Price", "Floor Volume"),
         col = c("blue", "red"), lty = c(1, 2))
}

ts_floor_vol_scatter(ts_cmeS, 'CME')
ts_with_floor_vol_plot(ts_cmeS, 'CME')
```

### IMM

```{r, warning=F, message=F}
lm_imm <- lm(spline ~ floor.vol, data=ts_immS)
summary(lm_imm)

ts_floor_vol_scatter(ts_immS, 'IMM')
ts_with_floor_vol_plot(ts_immS, 'IMM')
```

### IOM

```{r lm_iom}
lm_iom <- lm(spline ~ floor.vol, data=ts_iomS)
summary(lm_iom)


ts_floor_vol_scatter(ts_iomS, 'IOM')
ts_with_floor_vol_plot(ts_iomS, 'IOM')
```

## Linear Regression with ARIMA Errors

### CME

```{r, warning=F, message=F}
library(forecast)

lm_arima_cme <- auto.arima(ts_cmeS$spline, xreg=ts_cmeS$floor.vol)
summary(lm_arima_cme)
```

### IMM

```{r, warning=F, message=F}
lm_arima_imm <- auto.arima(ts_immS$spline, xreg=ts_immS$floor.vol)
summary(lm_arima_imm)
```

### IOM

```{r, warning=F, message=F}
lm_arima_iom <- auto.arima(ts_iomS$spline, xreg=ts_iomS$floor.vol)
summary(lm_arima_iom)
```

## Holt Winters

### CME

```{r, warning=F, message=F}
hw_cme <- HoltWinters(ts(ts_cmeS$spline, start=2001, frequency = 12))
c(hw_cme$alpha, hw_cme$beta, hw_cme$gamma)
```

### IMM

```{r, warning=F, message=F}
hw_imm <- HoltWinters(ts(ts_immS$spline, start=2001, frequency = 12))
c(hw_imm$alpha, hw_imm$beta, hw_imm$gamma)
```

### IOM 

```{r, warning=F, message=F}
hw_iom <- HoltWinters(ts(ts_iomS$spline, start=2001, frequency = 12))
c(hw_iom$alpha, hw_iom$beta, hw_iom$gamma)
```

## ARIMA

### CME

```{r, warning=F, message=F}
arima_cme <- auto.arima(ts_cmeS$spline, seasonal = F)
summary(arima_cme)
```

```{r, warning=F, message=F}
auto.arima(ts(ts_cmeS$spline, start=2001, frequency = 12), seasonal = F)
```

### IMM

```{r, warning=F, message=F}
arima_imm <- auto.arima(ts_immS$spline, seasonal = F)
summary(arima_imm)
```

### IOM 

```{r, warning=F, message=F}
arima_iom <- auto.arima(ts_iomS$spline, seasonal = F)
summary(arima_iom)
```

## Seasonal ARIMA 

Seasonality is monthly

### CME

```{r, warning=F, message=F}
auto.arima(ts(ts_cmeS$spline, start=2001, frequency = 12))
```

`auto.arima` won't automatically pick up a seasonal compoment on the CME seat
version even when putting it into a time series object. So I'll manually fit 
one.

First plot with a difference and a lagged seasonal (12 month) difference.

```{r, warning=F, message=F}
ts_cmeS$spline %>% 
  ts(start=2001, frequency = 12) %>% 
  diff(lag=12) %>%
  diff() %>%
  ggtsdisplay()
```

```{r, warning=F, message=F}
ts_cmeS$spline %>%
  ts(start=2001, frequency = 12) %>% 
  Arima(order=c(c(0,1,1)), seasonal=c(0,0,1)) %>%
  residuals() %>%
  ggtsdisplay()
```

A significant spike at lag 2 indicates an additional non-seasoanl term is needed.

```{r, warning=F, message=F}
sarima_cme <- ts_cmeS$spline %>%
  ts(start=2001, frequency = 12) %>% 
  Arima(order=c(c(0,1,2)), seasonal=c(0,0,1))

checkresiduals(sarima_cme)

summary(sarima_cme)
```

That's okay enough I think. It's not great but it's okay.

### IMM

```{r, warning=F, message=F}
auto.arima(ts(ts_immS$spline, start=2001, frequency = 12))
```

Ugh not this again. Okay do the same thing as with CME. The ARIMA for this 
model was of the order c(0,1,1) so start with that plus some seasonality terms.

```{r, warning=F, message=F}
ts_immS$spline %>%
  ts(start=2001, frequency = 12) %>% 
  Arima(order=c(c(0,1,1)), seasonal=c(0,0,1)) %>%
  residuals() %>%
  ggtsdisplay()
```

```{r, warning=F, message=F}
sarima_imm <- ts_immS$spline %>%
  ts(start=2001, frequency = 12) %>% 
  Arima(order=c(c(0,1,1)), seasonal=c(0,0,1))

checkresiduals(sarima_imm)

summary(sarima_imm)
```

That's fine. This model only needed a c(0,1,1) non-seasonal model where the CME
series needed a c(0,1,2). I think that's interesting.

### IOM 

```{r, warning=F, message=F}
auto.arima(ts(ts_iomS$spline, start=2001, frequency = 12))
```

Swell. This picked up a seaonality term all its own. We'll go with that.

```{r, warning=F, message=F}
sarima_iom <- ts_iomS$spline %>%
  ts(start=2001, frequency = 12) %>% 
  Arima(order=c(c(0,1,1)), seasonal=c(0,0,1))

checkresiduals(sarima_iom)

summary(sarima_iom)
```

## Fractional ARIMA

Check applicability first using the ACF. If the ACF decays very slowly
then the series could be a good candidate for an ARFIMA model. ARFIMA models
are good for detecting long-memory, and lots of autocorrelation at deep
lags mean that the series has a long memory, or depenence on lots of lagged
values.

### CME

Simulate a fractional ARIMA and plot the ACF of it against he ACF of the CME
time series so that it is easy to see if they are similar or dissimilar.

```{r, warning=F, message=F}
library(fracdiff)

ts.test <- fracdiff.sim( 1000, ar = .55, ma = -.4, d = .3) 

par(mfrow=c(2,1))
acf(ts.test$series, lag=50)
acf(ts_cmeS$spline, lag=50)
```

Looks to me like a slowly decaying ACF plot. Next, fit the ARFIMA.

```{r, warning=F, message=F}
arfima_cme <- arfima(ts(ts_cmeS$spline, start=2001, frequency = 12))

summary(arfima_cme)
```

### IMM

```{r, warning=F, message=F}
acf(ts_immS$spline, lag=50)
```

```{r, warning=F, message=F}
arfima_imm <- arfima(ts_immS$spline)

summary(arfima_imm)
```

### IOM 

```{r, warning=F, message=F}
acf(ts_iomS$spline, lag=50)
```

```{r, warning=F, message=F}
arfima_iom <- arfima(ts_iomS$spline)

summary(arfima_iom)
```

## ARMA and GARCH combination 


GARCH models are used when a time series displays heteroskedasticity, that is,
non-constant variance. Specifically when the variance of the errors is 
autocorrelated.

Use the fGarch R library and garchFit()


### CME

```{r, warning=F, message=F}
acf(ts_cmeS$spline)
acf(ts_cmeS$spline^2)
```

While the ACF of the time series shows autocorrelation, the ACF of the 
squared time series shows that the variance might also have autocorrelation.

Visual inspection of the residuals of a fitted ARIMA model shows the
same thing.

```{r, warning=F, message=F}
plot(residuals(arima_cme), type='p')
```

The "accordian" pattern we discussed in class is clear here. The variance in
the residuals of an ARIMA model is not constant, but varies, and looks to
be autocorrelated (maybe with an intervention).

### Fitting the God Damn GARCH

Nothing is working right now. I've tried garchFit from fGarch and 
ugarchfit from rugarch. I've tried different solvers and I've tried winsorizing
(removing outliers) from the dataset. I keep getting failed to converge or
failed to invert hessian errors.

```{r, warning=F, message=F}
library(fGarch)
library(TSA)

garch_cme <- garchFit(~arma(1,1)+garch(1,1), 
                      data=ts(ts_cmeS$spline, frequency = 12), 
                      trace=F)
summary(garch_cme)
McLeod.Li.test(y=residuals(garch_cme))
```

### IMM

```{r, warning=F, message=F}
# garch_imm <- garchFit(~arma(0,1)+garch(1,1), 
#                       data=ts(ts_immS$spline, frequency = 12), 
#                       trace=F)
# summary(garch_imm)
# McLeod.Li.test(y=residuals(garch_imm))
```

### IOM 

```{r, warning=F, message=F}
# garch_iom <- garchFit(~arma(0,1)+garch(1,1), 
#                       data=ts(ts_iomS$spline, frequency = 12), 
#                       trace=F)
# summary(garch_iom)
# McLeod.Li.test(y=residuals(garch_iom))
```

# Task B: Evaluation

Evaluated each model with sMAPE

```{r, warning=F, message=F}
smape <- function(act, pred) {
  sm <- abs(act - pred) / ((abs(act) + abs(pred))/2)
  return(sum(sm, na.rm = T) * 100/length(act))
}

forecast_2013_plot <- function(df_, col, division, model) {
  SMAPE <- smape(df_$spline, df_[,col])
  
  df_ %>%
    melt(id.vars='yearmon', measure.vars = c('spline', col)) %>%
    ggplot(aes(x=yearmon, y=value, color=variable)) + 
    geom_line() +
    scale_x_yearmon() + 
    labs(title=paste(model, 'Forecast for 2013', division, 'Seat Price'),
         subtitle=paste0('SMAPE: ', round(SMAPE, 1), '%'))
}
```

## Linear Regression

```{r, warning=F, message=F}
ts_cmeS_2013$lm_forecast <- predict(lm_cme, newdata=ts_cmeS_2013)
forecast_2013_plot(ts_cmeS_2013, 'lm_forecast', 'CME', 'Linear Regression')

ts_immS_2013$lm_forecast <- predict(lm_imm, newdata=ts_immS_2013)
forecast_2013_plot(ts_immS_2013, 'lm_forecast', 'IMM', 'Linear Regression')

ts_iomS_2013$lm_forecast <- predict(lm_iom, newdata=ts_iomS_2013)
forecast_2013_plot(ts_iomS_2013, 'lm_forecast', 'IOM', 'Linear Regression')

smape_df <- data.frame(
  row.names = c('cme', 'imm', 'iom'),
  Linear.Regression = c(with(ts_iomS_2013, smape(spline, lm_forecast)),
                        with(ts_immS_2013, smape(spline, lm_forecast)),
                        with(ts_cmeS_2013, smape(spline, lm_forecast))))
```

## Regression with ARIMA Errors

```{r, warning=F, message=F}
ts_cmeS_2013$lm_arima_forecast <- predict(lm_arima_cme, newxreg=ts_cmeS_2013$floor.vol)$pred
forecast_2013_plot(ts_cmeS_2013, 'lm_arima_forecast', 'CME', 'ARIMA with Regression Errors')

ts_immS_2013$lm_arima_forecast <- predict(lm_arima_imm, newxreg=ts_immS_2013$floor.vol)$pred
forecast_2013_plot(ts_immS_2013, 'lm_arima_forecast', 'IMM', 'ARIMA with Regression Errors')

ts_iomS_2013$lm_arima_forecast <- predict(lm_arima_iom, newxreg=ts_iomS_2013$floor.vol)$pred
forecast_2013_plot(ts_iomS_2013, 'lm_arima_forecast', 'IOM', 'ARIMA with Regression Errors')

smape_df$LM.ARIMA <- c(
  with(ts_cmeS_2013, smape(spline, lm_arima_forecast)),
  with(ts_immS_2013, smape(spline, lm_arima_forecast)),
  with(ts_iomS_2013, smape(spline, lm_arima_forecast))
)
```

## Holt-Winters

```{r, warning=F, message=F}
ts_cmeS_2013$hw_forecast <- forecast(hw_cme, h=12)$mean
forecast_2013_plot(ts_cmeS_2013, 'hw_forecast', 'CME', 'Holt-Winters')

ts_immS_2013$hw_forecast <- forecast(hw_imm, h=12)$mean
forecast_2013_plot(ts_immS_2013, 'hw_forecast', 'IMM', 'Holt-Winters')

ts_iomS_2013$hw_forecast <- forecast(hw_iom, h=12)$mean
forecast_2013_plot(ts_iomS_2013, 'hw_forecast', 'IOM', 'Holt-Winters')

smape_df$Holt.Winters <- c(
  with(ts_cmeS_2013, smape(spline, hw_forecast)),
  with(ts_immS_2013, smape(spline, hw_forecast)),
  with(ts_iomS_2013, smape(spline, hw_forecast))
)
```

## ARIMA

```{r, message=F, warning=F}
arima_order <- function(fit_) {
  order <- setNames(fit_$arma, c("p", "q", "P", "Q", "m", "d", "D"))
  pdq <- order[c('p','d','q')]
  PDQ <- order[c('P','D','Q')]
  
  name <- paste0("ARIMA(", paste(pdq, collapse=","), ")")
  
  if(sum(PDQ) > 0) {
    name <- paste0(name, "(", paste(PDQ, collapse=","), ")", "[", order['m'], "]")
  }
  return(name)
}


ts_cmeS_2013$arima_forecast <- forecast(arima_cme, h=12)$mean
forecast_2013_plot(ts_cmeS_2013, 'arima_forecast', 'CME', arima_order(arima_cme))

ts_immS_2013$arima_forecast <- forecast(arima_imm, h=12)$mean
forecast_2013_plot(ts_immS_2013, 'arima_forecast', 'IMM', arima_order(arima_imm))

ts_iomS_2013$arima_forecast <- forecast(arima_iom, h=12)$mean
forecast_2013_plot(ts_iomS_2013, 'arima_forecast', 'IOM', arima_order(arima_iom))

smape_df$ARIMA <- c(
  with(ts_cmeS_2013, smape(spline, arima_forecast)),
  with(ts_immS_2013, smape(spline, arima_forecast)),
  with(ts_iomS_2013, smape(spline, arima_forecast))
)
```

## Seasonal ARIMA

```{r, warning=F, warning=F}
ts_cmeS_2013$sarima_forecast <- forecast(sarima_cme, h=12)$mean
forecast_2013_plot(ts_cmeS_2013, 'sarima_forecast', 'CME', arima_order(sarima_cme))

ts_immS_2013$sarima_forecast <- forecast(sarima_imm, h=12)$mean
forecast_2013_plot(ts_immS_2013, 'sarima_forecast', 'IMM', arima_order(sarima_imm))

ts_iomS_2013$sarima_forecast <- forecast(sarima_iom, h=12)$mean
forecast_2013_plot(ts_iomS_2013, 'sarima_forecast', 'IOM', arima_order(sarima_iom))

smape_df$SARIMA <- c(
  with(ts_cmeS_2013, smape(spline, sarima_forecast)),
  with(ts_immS_2013, smape(spline, sarima_forecast)),
  with(ts_iomS_2013, smape(spline, sarima_forecast))
)
```

## Fractional ARIMA

```{r, warning=F, message=F}
ts_cmeS_2013$arfima_forecast <- forecast(arfima_cme, h=12)$mean
forecast_2013_plot(ts_cmeS_2013, 'arfima_forecast', 'CME', 'ARFIMA')

ts_immS_2013$arfima_forecast <- forecast(arfima_imm, h=12)$mean
forecast_2013_plot(ts_immS_2013, 'arfima_forecast', 'IMM', 'ARFIMA')

ts_iomS_2013$arfima_forecast <- forecast(arfima_iom, h=12)$mean
forecast_2013_plot(ts_iomS_2013, 'arfima_forecast', 'IOM', 'ARFIMA')

smape_df$ARFIMA <- c(
  with(ts_cmeS_2013, smape(spline, arfima_forecast)),
  with(ts_immS_2013, smape(spline, arfima_forecast)),
  with(ts_iomS_2013, smape(spline, arfima_forecast))
)
```

## Garch

```{r, warning=F, message=F}
ts_cmeS_2013$garch_forecast <- predict(garch_cme, 12)$meanForecast
forecast_2013_plot(ts_cmeS_2013, 'garch_forecast', 'CME', 'GARCH')

#ts_cmeS_2013$garch_forecast <- predict(garch_imm, 12)$meanForecast
#forecast_2013_plot(ts_immS_2013, 'garch_forecast', 'IMM', 'GARCH')

#ts_cmeS_2013$garch_forecast <- predict(garch_iom, 12)$meanForecast
#forecast_2013_plot(ts_iomS_2013, 'garch_forecast', 'IOM', 'GARCH')

smape_df$GARCH <- c(
  with(ts_cmeS_2013, smape(spline, garch_forecast)),
  NA,#with(ts_immS_2013, smape(spline, arfima_forecast))
  NA #with(ts_iomS_2013, smape(spline, arfima_forecast))
)
```

# Recommendation

```{r, warning=F, message=F}
data.frame(t(smape_df))
```

I would recommend seasonal ARIMA. While ARIMA did slighly better on the 
SMAPE metric, it just predicted point forecasts which is silly.
