---
title: "31006: Assignment 4"
subtitle: "Seasonality with Traffic Flow"
author: "Scott Shepard"
date: "4/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prompt

The daily data is from Illinois Dept of Transporation (IDOT) for I80E 1EXIT 
(the 2nd data column) - note each data point is an hourly count of the number 
of vehicles at a specific location on I80E.

Use the daily data for last 2 weeks of June 2013 to develop an 
ARIMA forecasting model.

Objective is to forecast the hourly counts for July 1.

The actual data file for July 1 is included for you to test your estimate.

**Part 1** - Use ARIMA(p,d,q) model to forecast. 
Find the model returned by R `auto.arima()`. 
Change the values of p and q and determine the best model using AICc and BIC. 
Do AICc and BIC select the same model as the best model?

**Part 2** - Use day of the week seasonal ARIMA(p,d,q)(P,Q,D)s model to 
forecast for July 1 (which is a Monday) - note use the hourly data

**Part 3** - Use hour of the day seasonal ARIMA (p,d,q)(P,D,Q)s model to 
forecast for the hours 8:00, 9:00, 17:00 and 18:00 on July 1

**Part 4** - For the July 1 8:00, 9:00, 17:00 and 18:00 forecasts, 
which model is better (part 2 or part 3) ?

## Part 0: Reading Data 

```{r data, warning=FALSE, message=FALSE}
library(gdata)
library(stringr)

base_path <- "~/Datasets/31006/1-57 Traffic/"

date_parse_helper <- function(d) {
  d <- str_replace_all(d, "I-57-|\\.xls", "")
  as.Date(d, format = "%Y-%B-%d")
}

read_i57_xls <- function(base_path, file_name) {
  df_ <- read.xls(file.path(base_path, file_name), skip=4, header=F, stringsAsFactors=F)
  df_$date <- date_parse_helper(file_name)
  df_$datetime <- as.POSIXct(paste(df_$date, df_[,3]), format="%Y-%m-%d %H:%M")
  df_[1:24,]
}

df <- plyr::ldply(list.files(base_path)[c(2:16,1)], function(f) read_i57_xls(base_path, f))
head(df[,1:10])
```

The field we are trying to forecast here is V5.

```{r, message=FALSE}
library(xts)

time_index <- seq(from = as.POSIXct("2013-06-16 01:00"), 
                  to = as.POSIXct("2013-07-02 00:00"), by = "hour")

obs <- xts(df[,5], order.by = time_index, frequency = 24)

train <- obs['2013-06-16/2013-06-30']
test <- obs['2013-07-01']

plot(obs)
```

Data looks highly seasonal, a strong time-based component. Traffic is lowest
around 2am, with a double peak during weekdays aroung 9am and 5pm, and a
single peak on weekdays at midday. Visually, it looks like the hourly 
compoment will be more important than the weekday component.

## Part 1: ARIMA

Use ARIMA(p,d,q) model to forecast. Find the model returned by R `auto.arima()`. 
Change the values of p and q and determine the best model using AICc and BIC. 
Do AICc and BIC select the same model as the best model?

```{r}
library(forecast)

m0 <- auto.arima(train)
m1 <- Arima(train, c(4,0,6))
m2 <- Arima(train, c(8,0,12))

compare_models <- data.frame(m0[c('aic', 'bic', 'loglik')])
names(compare_models) <- c('aic', 'bic', 'loglik')
compare_models <- rbind(compare_models, data.frame(m1[c('aic', 'bic', 'loglik')]))
compare_models <- rbind(compare_models, data.frame(m2[c('aic', 'bic', 'loglik')]))
compare_models$model <- c('m0','m1','m2')
compare_models
```

Using AIC and BIC alone do not give the same model as using `auto.arima`. AIC 
continues to fall as more and more terms are added. BIC penalized additional
terms and so increases again when the terms get too large, however even BIC
prefered values of p and q higher than the "best" model. Even rerunning
`auto.arima` with higher max p and q values yields the same answer.

## Part 2: Day of Week Seasonality

Use day of the week seasonal ARIMA(p,d,q)(P,Q,D)s model to forecast for July 1 
(which is a Monday) - note use the hourly data

```{r}
dow_arima <- auto.arima(ts(train, frequency = 168))
dow_forecast <- forecast(dow_arima, h=24)
dow_forecast
```


```{r}
hour_arima <- auto.arima(ts(train, frequency = 24))
hourly_forecast <- forecast(hour_arima, h=24)
hourly_forecast
```


```{r}
matplot(index(test), cbind(test, data.frame(hourly_forecast)$Point.Forecast, data.frame(dow_forecast)$Point.Forecast))
```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```

