---
title: "31006: Assignment2"
author: "Scott Shepard"
date: "4/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Istanbul Timeseries Exploration

Prompt:
1. Determine if all the TS are stationary 
   a. qualitatively
   b. quantitatively
2. Linearly regress ISE against the remaining 5 stock index returns - determine which coefficients are significant
3. For the non-significant coefficients, continue to lag by 1 day until all coefficients are significant - how many lags are needed?
4. Concept question - what do you find in #1 and why?

```{r packages}
library(zoo)
library(xts)
suppressMessages(library(tseries))
```

```{r data}
df <- read.csv("~/Datasets/TS regression - data_akbilgic.csv")
df$date <- as.Date(as.character(df$date), format='%d-%b-%y')

ISE  <- xts(df$ISE, df$date)
SP   <- xts(df$SP,  df$date)
DAX  <- xts(df$DAX, df$date)
FTSE <- xts(df$FTSE, df$date)
NIKKEI <- xts(df$NIKKEI, df$date)
BOVESPA <- xts(df$BOVESPA, df$date)
```

## 1a. Determine if all the TS are stationary - Qualitative

The first step to both data exploration and determing if a time series is
stationary is to plot it. I'm plotting all six time series here. A stationary
time series will have an unchanging mean and should look stable, that is with
no upward or downward trend.

```{r ts_plots}
par(mfrow=c(3,2))
plot(ISE)
plot(SP)
plot(DAX)
plot(FTSE)
plot(NIKKEI)
plot(BOVESPA)
```

Visually all of these time series look stationary. None of them have an 
obvious trend. They are all more or less centered at 0 (not important for the
mean to be 0 but it is nice). I would say, qualitatively at least, that these
time series are stationary.

## 1.b Determine of all the TS are stationary -- quantitatively

One tool in our disposal for determining is a time series is stationary is the 
`kpss.test()` function. The KPSS test tests against the alternative that the 
series has a unit root. If a series has a unit root, it is has a stochastic 
trend that is unpredictable. In the KPSS test, if the p-value is large then the
null is accepted. Since the unit root is the alternative hypothesis, a low p-value
means that the time series is stationary, or at least there isn't enough evidence
to conclude that it is not stationary.

```{r}
library(tseries)

kpss.test(ISE)
kpss.test(SP)
kpss.test(DAX)
kpss.test(FTSE)
kpss.test(NIKKEI)
kpss.test(BOVESPA)
```

None of the KPSs levels are large enough to reject the null hypothesis,  
i.e. all these trends are stationary.


## 2. Linear Regression

Linearly regress ISE against remianing 5 stock index returns and determin
which coefficients are significant.

```{r}
fit <- lm(ISE ~ ., data=df[,-1])

summary(fit)
```

DAX, FTSE, and NIKKEI are significant coefficients. BOVESPA is significant 
at a 0.1 level so we won't count that one. SP is not significant at all.

About half the variance is explained by the regression model.

Explore the fitted model a bit.

```{r}
library(dplyr)
library(reshape2)
library(ggplot2)

df$ISE_fitted <- y_hat <- predict(fit)

melt(df, id.vars=('date'), measure.vars=c('ISE', 'ISE_fitted')) %>% 
  ggplot(aes(x=date, y=value, color=variable)) + geom_point()

ggplot(df, aes(x=ISE, y=ISE_fitted)) + geom_point()

par(mfrow=c(2,1))
plot(fit$residuals)

hist(fit$residuals)
```

The fit is decent. The residuals show no pattern and are normally distributed.
Plotting the real vs fitted values shows a very clear linear pattern but the 
overall scatter shows a lot of variance that is not captured. 

## 3. Lag Non-Significant Coefficients

We can improve the model by lagging the non-significant coefficients until 
they become sigfificant.

```{r}
library(DataCombine)

lag = 1
df_slide = slide(df, Var='SP', TimeVar='date', NewVar=paste0('SP_', lag), slideBy=-lag)
drops = c('date', 'ISE_fitted', 'SP')
summary(lm(ISE ~ ., data=df_slide[,!names(df_slide) %in% drops]))
```

A single day lag on SP brings the two non-significant coefficients into 
significance. I tested other combinations that also worked, like two days on SP
and one day on BOVESPA, but simplier is usually better when two models achieve
the same results.

## 4. What do you find in #1 and why?

In part 1 I found that all the timeseries were stationary. They are return 
indexes which means it's the daily return from open to close (or maybe from 
close the previous day). I'm rather surprised that they are all stationary 
as there should be a definite upward trend in daily return from 2009-2010. 
Maybe this is already trend corrected data?
