---
title: "Assignment6"
author: "Scott Shepard"
date: "5/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prompt

Input gas rate - this is the independent variable

Output gas CO2 % - this is the dependent variable that needs to be forecast

Tasks:

1. Use linear regression model - plot the ACF - what can you conclude ?
2. Use ARIMA (0,0,1) model for the residuals. Adjust the Input gas rate and Output CO2 % with the MA coefficient. Combine with the linear regression model. Plot the residuals.
3. Use ARIMA (1,0,0) model for the residuals. Adjust the Input gas rate and Output CO2 % with the AR coefficient. Combine with the linear regression model. Plot the residuals.
4. Use ARIMA (0,0,2) model for the residuals. Adjust the Input gas rate and Output CO2 % with the MA coefficient. Combine with the linear regression model. Plot the residuals.
5. Use ARIMA (2,0,0) model for the residuals. Adjust the Input gas rate and Output CO2 % with the AR coefficient. Combine with the linear regression model. Plot the residuals.
6. Use ARIMA (2,0,2) model for the residuals. Adjust the Input gas rate and Output CO2 % with the AR and MA coefficients. Combine with the linear regression model. Plot the residuals.
7. Use fractional ARIMA model (aka ARFIMA) for the output gas CO2% - plot the residuals, acf and pacf plots of the model. You can use an R package like fracdiff – be careful to determine which lag to choose when executing this test.
8. Perform  Summaries,  Durbin-Watson and Box-Ljung tests for each model and build table to compare AICs, BICs and p-vaules for each test across the ARIMA and ARFIMA models.
9. Based on ACF plots and test results, which ARIMA model gives the best result in terms of residuals being close to white noise ?

## 1. Linear Regression

```{r}
df <- read.csv('~/Datasets/31006/Gas Furnace Dataset.csv')
names(df) <- c('input', 'output')
head(df)
```


```{r}
m <- lm(output ~ input, data=df)

acf(m$residuals)
```

Serious autocorrelation of the residuals at low lags.

```{r}
plot(m$residuals, type='l', main='Residuals of Linear Model')
```

It's very clear that there is information that the linear model is not 
capturing.

## 2. ARIMA (0,0,1) 

```{r, message=FALSE}
library(forecast)

ma1 <- Arima(m$residuals, c(0,0,1))
```

## 3. ARIMA (1,0,0)

```{r}

```

## 4. ARIMA (0,0,2)

```{r}

```

## 5. ARIMA (2,0,0)

```{r}

```

## 6. ARIMA (2,0,2)

```{r}

```

## 7. Fractional ARIMA

```{r}

```

## 8. Compare Models

```{r}

```

## 9. Conclusions

```{r}
arma.innovation <- function(x, arma.model, ar.truncation=10) {
  p <- arma.model$arma[1]
  q <- arma.model$arma[2]
  ar.coef <- arma.model$coef[seq_len(p)]
  ma.coef <- arma.model$coef[p + seq_len(q)]
  if (q == 0) {
    infinite.ar.coef <- ar.coef
  } else {
    infinite.ar.coef <- -ARMAtoMA(-ma.coef, -ar.coef, ar.truncation)
  }
  return(as.vector(filter(x, c(1, -infinite.ar.coef), side=1)))
}
```
