---
title: "Assignment 8"
author: "Scott Shepard"
date: "6/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prompt

Task A:

Use the following algorithms:

- Linear regression (seat price is independent, volume(s) dependent)
- Linear regression with ARMA errors (use arima with xreg)
- Holt Winters
- ARIMA
- Seasonal ARIMA (SARIMA) - here seasonality is monthly
- Fractional ARIMA (ARFIMA) - check applicability first using the ACF
- ARMA and GARCH combination - use the fGarch R library and garchFit()

Note that you have to implement each of the above algorithms for each of the 3 classes of seats: CME, IMM, IOM. 

Task B:

Evaluate each algorithm from Task A using sMAPE. 
Which one do you recommend to forecast monthly prices for each of the seat classes?

# Data

## Create Time Serieses

This function is copied over from assignment 7.

```{r, warning=F, message=F}
cmeS <- read.csv("~/Datasets/31006/cmeS.csv")
immS <- read.csv("~/Datasets/31006/immS.csv")
iomS <- read.csv("~/Datasets/31006/iomS.csv")

library(dplyr)
library(zoo)
library(ggplot2)
library(reshape2)

create_timeseries <- function(df_, interpolation='spline') {
  # Date formatting
  df_$DateOfSale <- as.Date(df_$DateOfSale, '%m/%d/%Y')
  df_$yearmon <- as.yearmon(df_$DateOfSale)
  
  # Group by month
  df_ <- df_ %>%
    group_by(yearmon) %>%
    summarize(price=mean(price)) %>%
    data.frame() %>%
    merge(data.frame(yearmon = as.yearmon(2001 + seq(0, 12*13-1)/12)),all=T)
  
  # Interpolate
  df_$interpolated <- is.na(df_$price)
  x = df_$yearmon
  if('constant' %in% interpolation) {
    df_$constant <- approx(df_$yearmon, df_$price, xout=x, method = "constant")$y  
  }
  if('linear' %in% interpolation) {
    df_$linear <- approx(df_$yearmon, df_$price, xout=x)$y
  }
  if('spline' %in% interpolation) {
    df_$spline <- spline(df_, n=nrow(df_))$y
  }
  df_
}

cmeS <- create_timeseries(cmeS)
immS <- create_timeseries(immS)
iomS <- create_timeseries(iomS)
```

## Contract Volume Data

The new datasets need to be merged in with the time series.
  1. Contracts Volume
  2. Contracts Classification

```{r, message=FALSE, echo=FALSE}
library(stringr)

cv <- read.csv("~/Datasets/31006/Contracts_Volume.csv", stringsAsFactors = F)
cc <- read.csv("~/Datasets/31006/Contracts_Classification.csv", stringsAsFactors = F)

cv$Electronic.Volume <- str_replace_all(cv$Electronic.Volume, ',', '')
cv$Electronic.Volume <- as.numeric(cv$Electronic.Volume)
cv$Floor.Volume <- cv$Total.Volume - cv$Electronic.Volume

df = merge(cv, cc, by.x='Commodity.Indicator', by.y='Commodity.Code', all.x=T)

df$Date <- as.Date(df$Date, '%m/%d/%Y')
df$yearmon <- as.yearmon(df$Date)

# IMM and IOM can only trade the commodities listed under those seats
imm_iom_monthly_vol <- df %>% 
  dplyr::filter(Division %in% c('IMM', 'IOM')) %>%
  group_by(yearmon, Division) %>%
  summarize(electric.vol = sum(Electronic.Volume), 
            total.vol = sum(Total.Volume), 
            floor.vol=sum(Floor.Volume))

# CME can trade all commodities, not just the ones listed
cme_monthly_vol <- df %>%
  group_by(yearmon) %>%
  summarize(electric.vol = sum(Electronic.Volume), 
            total.vol = sum(Total.Volume), 
            floor.vol=sum(Floor.Volume)) %>%
  mutate(Division = 'CME')

monthly_vol <- rbind(data.frame(cme_monthly_vol), data.frame(imm_iom_monthly_vol))

ggplot(monthly_vol, aes(x=yearmon, y=floor.vol, color=Division)) + 
  geom_line() + 
  labs(title='Monthly Floor Volume at the Merch',
       y='Floor Volume', x='Time')
```

```{r}
attach_monthly_vol <- function(ts_df, division, monthly_df = monthly_vol) {
  monthly_df <- dplyr::filter(monthly_df, Division==division)
  monthly_df <- select(monthly_df, -Division)
  left_join(ts_df, monthly_df, by='yearmon')
}

ts_cmeS_all <- attach_monthly_vol(cmeS, 'CME')
ts_immS_all <- attach_monthly_vol(immS, 'IMM')
ts_iomS_all <- attach_monthly_vol(iomS, 'IOM')
```

### Train / Test

Before fitting any models, remove the 2013 year from the timeseries objects
and keep the 2013 year as the values to judge the forecasts.

```{r}
ts_cmeS_2013 <- dplyr::filter(ts_cmeS_all, yearmon >= 2013)
ts_immS_2013 <- dplyr::filter(ts_immS_all, yearmon >= 2013)
ts_iomS_2013 <- dplyr::filter(ts_iomS_all, yearmon >= 2013)

ts_cmeS <- dplyr::filter(ts_cmeS_all, yearmon < 2013)
ts_immS <- dplyr::filter(ts_immS_all, yearmon < 2013)
ts_iomS <- dplyr::filter(ts_iomS_all, yearmon < 2013)
```

# Task A: Fit Models

## Linear Regression 

Seat price is independent, Volume dependent

### CME

```{r}
ts_with_floor_vol_plot <- function(ts_df, division) {
  X <- ts_df$yearmon
  Y1 <- ts_df$spline
  Y2 <- ts_df$floor.vol
  par(mar = c(5, 5, 3, 5))
  plot(x=X, y=Y1, type ="l", 
       ylab = paste("Interpolated Price of",division,"Seat"),
       main = paste(division, "Seat price & Floor Volume"), 
       xlab = "Time", xaxt='n',
       col = "blue")
  par(new = TRUE)
  plot(x=X, y=Y2, type = "l", 
       xaxt = "n", yaxt = "n", ylab = "", xlab = "", 
       col = "red", lty = 2)
  axis(side = 4)
  mtext("Floor Volume", side = 4, line = 3)
  legend("topleft", c("Seat Price", "Floor Volume"),
         col = c("blue", "red"), lty = c(1, 2))
}

lm_cme <- lm(spline ~ floor.vol, data=ts_cmeS)
ts_with_floor_vol_plot(ts_cmeS, 'CME')
```

### IMM

I won't print the other summaries or plots for simplicity.

```{r}
lm_imm <- lm(spline ~ floor.vol, data=ts_immS)
```

### IOM

```{r lm_iom}
lm_iom <- lm(spline ~ floor.vol, data=ts_iomS)
```

## Linear Regression with ARIMA Errors

### CME

```{r, warning=F, message=F}
library(forecast)

lm_arima_cme <- auto.arima(ts_cmeS$spline, xreg=ts_cmeS$floor.vol)
summary(lm_arima_cme)
```

### IMM

```{r}
lm_arima_imm <- auto.arima(ts_immS$spline, xreg=ts_immS$floor.vol)
```

### IOM

```{r}
lm_arima_iom <- auto.arima(ts_iomS$spline, xreg=ts_iomS$floor.vol)
```

## Holt Winters

### CME

```{r}
hw_cme <- HoltWinters(ts(ts_cmeS$spline, start=2001, frequency = 12))
c(hw_cme$alpha, hw_cme$beta, hw_cme$gamma)
```

### IMM

```{r}
hw_imm <- HoltWinters(ts(ts_immS$spline, start=2001, frequency = 12))
c(hw_imm$alpha, hw_imm$beta, hw_imm$gamma)
```

### IOM 

```{r}
hw_iom <- HoltWinters(ts(ts_iomS$spline, start=2001, frequency = 12))
c(hw_iom$alpha, hw_iom$beta, hw_iom$gamma)
```

## ARIMA

### CME

```{r}
arima_cme <- auto.arima(ts_cmeS$spline, seasonal = F)
summary(arima_cme)
```

```{r}
auto.arima(ts(ts_cmeS$spline, start=2001, frequency = 12), seasonal = F)
```

### IMM

```{r}
arima_imm <- auto.arima(ts_immS$spline, seasonal = F)
```

### IOM 

```{r}
arima_iom <- auto.arima(ts_iomS$spline, seasonal = F)
summary(arima_iom)
```

## Seasonal ARIMA 

Seasonality is monthly

### CME

```{r}
auto.arima(ts(ts_cmeS$spline, start=2001, frequency = 12))
```

`auto.arima` won't automatically pick up a seasonal compoment on the CME seat
version even when putting it into a time series object. So I'll manually fit 
one. 

I looged at the acf or plots with a lagged 12 month seasonal difference and
added terms to the ARMA part accordingly until I found (0,1,2)(0,0,1)[12]

First plot with a difference and a lagged seasonal (12 month) difference.

```{r}
sarima_cme <- ts_cmeS$spline %>%
  ts(start=2001, frequency = 12) %>% 
  Arima(order=c(c(0,1,2)), seasonal=c(0,0,1))

checkresiduals(sarima_cme)

summary(sarima_cme)
```

That's okay I think. It's not great but it's okay.

### IMM

IMM also didn't pick up a seasonality component from auto.arima so 
I again manually fit one. This time the results were slightly different.

```{r}
sarima_imm <- ts_immS$spline %>%
  ts(start=2001, frequency = 12) %>% 
  Arima(order=c(c(0,1,1)), seasonal=c(0,0,1))

checkresiduals(sarima_imm)

summary(sarima_imm)
```

That's fine. This model only needed a c(0,1,1) non-seasonal model where the CME
series needed a c(0,1,2). I think that's interesting.

### IOM 

For this series, auto.arima picked up a seaonality term all its own. 
We'll go with that.

```{r}
sarima_iom <- ts_iomS$spline %>%
  ts(start=2001, frequency = 12) %>% 
  Arima(order=c(c(0,1,1)), seasonal=c(0,0,1))
```

## Fractional ARIMA

Check applicability first using the ACF. If the ACF decays very slowly
then the series could be a good candidate for an ARFIMA model. ARFIMA models
are good for detecting long-memory, and lots of autocorrelation at deep
lags mean that the series has a long memory, or depenence on lots of lagged
values.

### CME

Simulate a fractional ARIMA and plot the ACF of it against he ACF of the CME
time series so that it is easy to see if they are similar or dissimilar.

```{r}
library(fracdiff)

ts.test <- fracdiff.sim( 1000, ar = .55, ma = -.4, d = .3) 

par(mfrow=c(2,1))
acf(ts.test$series, lag=50)
acf(ts_cmeS$spline, lag=50)
```

Looks to me like a slowly decaying ACF plot. Next, fit the ARFIMA.

```{r}
arfima_cme <- arfima(ts(ts_cmeS$spline, start=2001, frequency = 12))

summary(arfima_cme)
```

### IMM

```{r}
acf(ts_immS$spline, lag=50)
```

```{r}
arfima_imm <- arfima(ts_immS$spline)

summary(arfima_imm)
```

### IOM 

```{r}
acf(ts_iomS$spline, lag=50)
```

```{r}
arfima_iom <- arfima(ts_iomS$spline)

summary(arfima_iom)
```

## ARMA and GARCH combination 

GARCH models are used when a time series displays heteroskedasticity, 
that is, non-constant variance. 
Specifically when the variance of the errors is 
autocorrelated.

Use the fGarch R library and garchFit()

### CME

```{r, warning=F, message=F}
acf(ts_cmeS$spline)
acf(ts_cmeS$spline^2)
```

While the ACF of the time series shows autocorrelation, the ACF of the 
squared time series shows that the variance might also have autocorrelation.

A visual inspection of the residuals would show the same thing, for 
brevity's sake I'll omit it here and go straight to model fitting.

I had to use an `arma(1,0)` model to get the `garchFit` to work even
though `auto.arima` did not detect an AR term when fitting that model
above. Otherwise the hessian failed to invert

```{r, warning=F, message=F}
library(fGarch)
library(TSA)

garch_cme <- garchFit(~arma(1,0)+garch(1,1), 
                      data=ts(ts_cmeS$spline, frequency = 12), 
                      trace=F)
summary(garch_cme)
McLeod.Li.test(y=residuals(garch_cme))
```

### IMM

I cannot get garchFit to work on IMM or IOM series. It just
gives the error: 

```
Error in solve.default(fit$hessian) : 
  system is computationally singular: reciprocal condition number = 2.04539e-21
```

I have tried using the uGarchFit package, winsorizing the outliers,
and different solver but nothing seems to work. Not sure why it 
worked for CME but not IMM or IOM.


```{r}
# garch_imm <- garchFit(~arma(1,0)+garch(1,1), 
#                       data=ts(ts_immS$spline, frequency = 12), 
#                       trace=F)
# summary(garch_imm)
# McLeod.Li.test(y=residuals(garch_imm))
```

### IOM 

```{r}
# garch_iom <- garchFit(~arma(0,1)+garch(1,1), 
#                       data=ts(ts_iomS$spline, frequency = 12), 
#                       trace=F)
# summary(garch_iom)
# McLeod.Li.test(y=residuals(garch_iom))
```

# Task B: Evaluation

Evaluated each model with sMAPE.  
I'll only plot the forecasts for the CME models.

```{r}
smape <- function(act, pred) {
  sm <- abs(act - pred) / ((abs(act) + abs(pred))/2)
  return(sum(sm, na.rm = T) * 100/length(act))
}

forecast_2013_plot <- function(df_, col, division, model) {
  SMAPE <- smape(df_$spline, df_[,col])
  
  df_ %>%
    melt(id.vars='yearmon', measure.vars = c('spline', col)) %>%
    ggplot(aes(x=yearmon, y=value, color=variable)) + 
    geom_line() +
    scale_x_yearmon() + 
    labs(title=paste(model, 'Forecast for 2013', division, 'Seat Price'),
         subtitle=paste0('SMAPE: ', round(SMAPE, 1), '%'))
}
```

## Linear Regression

```{r,  warning=F, message=F}
ts_cmeS_2013$lm_forecast <- predict(lm_cme, newdata=ts_cmeS_2013)
ts_immS_2013$lm_forecast <- predict(lm_imm, newdata=ts_immS_2013)
ts_iomS_2013$lm_forecast <- predict(lm_iom, newdata=ts_iomS_2013)

smape_df <- data.frame(
  row.names = c('cme', 'imm', 'iom'),
  Linear.Regression = c(with(ts_iomS_2013, smape(spline, lm_forecast)),
                        with(ts_immS_2013, smape(spline, lm_forecast)),
                        with(ts_cmeS_2013, smape(spline, lm_forecast))))

forecast_2013_plot(ts_cmeS_2013, 'lm_forecast', 'CME', 'Linear Regression')
```

Not a great forecast.

## Regression with ARIMA Errors

```{r, warning=F, message=F}
ts_cmeS_2013$lm_arima_forecast <- 
  predict(lm_arima_cme, newxreg=ts_cmeS_2013$floor.vol)$pred

ts_immS_2013$lm_arima_forecast <- 
  predict(lm_arima_imm, newxreg=ts_immS_2013$floor.vol)$pred

ts_iomS_2013$lm_arima_forecast <- 
  predict(lm_arima_iom, newxreg=ts_iomS_2013$floor.vol)$pred

smape_df$LM.ARIMA <- c(
  with(ts_cmeS_2013, smape(spline, lm_arima_forecast)),
  with(ts_immS_2013, smape(spline, lm_arima_forecast)),
  with(ts_iomS_2013, smape(spline, lm_arima_forecast))
)

forecast_2013_plot(ts_cmeS_2013, 'lm_arima_forecast', 'CME', 'ARIMA with Regression Errors')
```

## Holt-Winters

```{r, warning=F, message=F}
ts_cmeS_2013$hw_forecast <- forecast(hw_cme, h=12)$mean
ts_immS_2013$hw_forecast <- forecast(hw_imm, h=12)$mean
ts_iomS_2013$hw_forecast <- forecast(hw_iom, h=12)$mean

smape_df$Holt.Winters <- c(
  with(ts_cmeS_2013, smape(spline, hw_forecast)),
  with(ts_immS_2013, smape(spline, hw_forecast)),
  with(ts_iomS_2013, smape(spline, hw_forecast))
)

forecast_2013_plot(ts_cmeS_2013, 'hw_forecast', 'CME', 'Holt-Winters')
```

## ARIMA

```{r, warning=F, message=F}
arima_order <- function(fit_) {
  order <- setNames(fit_$arma, c("p", "q", "P", "Q", "m", "d", "D"))
  pdq <- order[c('p','d','q')]
  PDQ <- order[c('P','D','Q')]
  
  name <- paste0("ARIMA(", paste(pdq, collapse=","), ")")
  
  if(sum(PDQ) > 0) {
    name <- paste0(name, "(", paste(PDQ, collapse=","), ")", "[", order['m'], "]")
  }
  return(name)
}

ts_cmeS_2013$arima_forecast <- forecast(arima_cme, h=12)$mean
ts_immS_2013$arima_forecast <- forecast(arima_imm, h=12)$mean
ts_iomS_2013$arima_forecast <- forecast(arima_iom, h=12)$mean

smape_df$ARIMA <- c(
  with(ts_cmeS_2013, smape(spline, arima_forecast)),
  with(ts_immS_2013, smape(spline, arima_forecast)),
  with(ts_iomS_2013, smape(spline, arima_forecast))
)

forecast_2013_plot(ts_cmeS_2013, 'arima_forecast', 'CME', arima_order(arima_cme))

```

## Seasonal ARIMA

```{r, warning=F, message=F}
ts_cmeS_2013$sarima_forecast <- forecast(sarima_cme, h=12)$mean
ts_immS_2013$sarima_forecast <- forecast(sarima_imm, h=12)$mean
ts_iomS_2013$sarima_forecast <- forecast(sarima_iom, h=12)$mean

smape_df$SARIMA <- c(
  with(ts_cmeS_2013, smape(spline, sarima_forecast)),
  with(ts_immS_2013, smape(spline, sarima_forecast)),
  with(ts_iomS_2013, smape(spline, sarima_forecast))
)

forecast_2013_plot(ts_cmeS_2013, 'sarima_forecast', 'CME', arima_order(sarima_cme))
```

## Fractional ARIMA

```{r, warning=F, message=F}
ts_cmeS_2013$arfima_forecast <- forecast(arfima_cme, h=12)$mean
ts_immS_2013$arfima_forecast <- forecast(arfima_imm, h=12)$mean
ts_iomS_2013$arfima_forecast <- forecast(arfima_iom, h=12)$mean

smape_df$ARFIMA <- c(
  with(ts_cmeS_2013, smape(spline, arfima_forecast)),
  with(ts_immS_2013, smape(spline, arfima_forecast)),
  with(ts_iomS_2013, smape(spline, arfima_forecast))
)

forecast_2013_plot(ts_cmeS_2013, 'arfima_forecast', 'CME', 'ARFIMA')
```

## Garch

```{r, warning=F, message=F}
ts_cmeS_2013$garch_forecast <- predict(garch_cme, 12)$meanForecast
#ts_cmeS_2013$garch_forecast <- predict(garch_imm, 12)$meanForecast
#ts_cmeS_2013$garch_forecast <- predict(garch_iom, 12)$meanForecast

smape_df$GARCH <- c(
  with(ts_cmeS_2013, smape(spline, garch_forecast)),
  NA,#with(ts_immS_2013, smape(spline, arfima_forecast))
  NA #with(ts_iomS_2013, smape(spline, arfima_forecast))
)

forecast_2013_plot(ts_cmeS_2013, 'garch_forecast', 'CME', 'GARCH')
```

# Recommendation

```{r}
data.frame(t(smape_df))
```

I would recommend seasonal ARIMA. While ARIMA did slighly better on the 
SMAPE metric, it just predicted point forecasts which is silly.
