---
title: "Assignment7"
author: "Scott Shepard"
date: "5/24/2019"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Prompt

The data contain seat prices from the Chicago Mercantile Exchange.

There are three classes of seats CME, IMM, and IOM which confer the rights to 
trade different sets of commodities traded. CME seat owners can trade 
everything, IMM seat owners are allowed to trade everything except 
agricultural products, and IOM seat owners are allowed to trade only in 
index products and all options.

The seat price files are marked as ***S. The files contain the price for 
which CME seats sold and the date of the sale. As expected, the seat price 
time series is irregularly spaced in time. 

Your task is to adopt an algorithm to create a time series that contains a 
seat price for each month starting from January 2001 to December 2013. 
You need to do this for the 3 classes of seats. Note that all 3 seat classes 
have sale prices for January 2001 so you should not have any start up issues. 
Please clearly explain why you adopted your algorithm and what other 
alternatives you may have considered and why you rejected the alternatives.

```{r}
cmeS <- read.csv("~/Datasets/31006/cmeS.csv")
immS <- read.csv("~/Datasets/31006/immS.csv")
iomS <- read.csv("~/Datasets/31006/iomS.csv")

head(cmeS)
```

To create a monthly timeseries from this data we need to:  
  * Average the data at the monthly level  
  * Interpolate when missing a month's data

```{r, warning=F, message=F}
library(dplyr)
library(zoo)
library(ggplot2)
library(reshape2)

create_timeseries <- function(df_, interpolation='spline') {
  # Date formatting
  df_$DateOfSale <- as.Date(df_$DateOfSale, '%m/%d/%Y')
  df_$yearmon <- as.yearmon(df_$DateOfSale)
  
  # Group by month
  df_ <- df_ %>%
    group_by(yearmon) %>%
    summarize(price=mean(price)) %>%
    data.frame() %>%
    merge(data.frame(yearmon = as.yearmon(2001 + seq(0, 12*13-1)/12)),all=T)
  
  # Interpolate
  df_$interpolated <- is.na(df_$price)
  x = df_$yearmon
  if('constant' %in% interpolation) {
    df_$constant <- approx(df_$yearmon, df_$price, xout=x, method = "constant")$y  
  }
  if('linear' %in% interpolation) {
    df_$linear <- approx(df_$yearmon, df_$price, xout=x)$y
  }
  if('spline' %in% interpolation) {
    df_$spline <- spline(df_, n=nrow(df_))$y
  }
  df_
}

df <- create_timeseries(cmeS, interpolation = c('constant', 'linear', 'spline'))
head(df)
```

We can now compare the interpolated points with three different methods:  
  * Constant: take previous value and populate foward  
  * Linear: Take first and last point and draw straight line  
  * Spline: Cublic spline function

```{r, warnings=F, message=F}
head(df, 24) %>%
  melt(id.vars = c('yearmon', 'interpolated')) %>%
  arrange(rev(variable)) %>%
  ggplot(aes(x=yearmon, y=value, color=variable, group=desc(variable))) + 
  geom_point() +
  geom_line() + 
  labs(title='Compare Interpolation Functions on Timeseries Data',
       subtitle='First 2 Years of CME Seats',
       x='Time', y='Seat Price')
```

It sure looks to me like the spline funtion does the best job at looking
like the timeseries. The constant and linear functions are very obvious
breaks in the series where the spline points could be real TS data if you 
didn't know.

Create the other timeseries

```{r}
ts_cmeS <- create_timeseries(cmeS)
ts_immS <- create_timeseries(immS)
ts_iomS <- create_timeseries(iomS)

ggplot(ts_cmeS, aes(x=yearmon, y=spline, color=interpolated)) + 
  geom_point() + scale_color_manual(values=c('black', 'red')) + 
  ggtitle('CME Seat Price: 2001-2013')

ggplot(ts_immS, aes(x=yearmon, y=spline, color=interpolated)) + 
  geom_point() + scale_color_manual(values=c('black', 'red')) +
  ggtitle('IMM Seat Price: 2001-2013')

ggplot(ts_iomS, aes(x=yearmon, y=spline, color=interpolated)) + 
  geom_point() + scale_color_manual(values=c('black', 'red')) + 
  ggtitle('IOM Seat Price: 2001-2013')
  
```




