---
title: "Assignment 5 - Multi Step Forecasts"
author: "Scott Shepard"
date: "5/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prompt

Load data from TSA package (the package is written by authors Jonathan Cryer and Kung-Sik Chan).
```
> library("TSA")
> data(beersales)
```
The data is the monthly beer sales in millions of barrels, 01/1975 - 12/1990.

Train: 01/1975 - 12/1989.

Test: 1990

**Part 1**

Use ARIMA(p,d,q) model to forecast beer sales for all months of 1990 using the following two multi-step forecasting approaches. For each model, check mean, autocorrelation and normality of the residuals. Confirm if the residuals are white noise.

*1A* - Use the h-period in forecast() to forecast each month of 1990. This is also known as recursive forecasting where you fit a model only once and use it recursively for h-periods.

*1B* - Use the monthly data as a continuous time series. Forecast for 1990 Jan, Plug forecast into the time series, build a new model to forecast for 1990 Feb. And so on and so forth. In other words, h=1 in all the forecasts. This is known as direct recursive (DirRec) forecasting where you fit a new model for each time step.

*1C* - Plot the mean, the p-value of the autocorrelation test and the p-value of the normality test of the residuals of the 12 models.

**Part 2**

Plot the Recursive and DirRec along with the actuals. Use ylim=c(12.5, 17) to get a good visual of the plot differences.

**Part 3**

Calculate the MSE for 1990 - which of the two approaches give better result and why?

# Part 1

Plot beersales data, split into training and test.

```{r, message=FALSE}
library(TSA)
library(xts)
library(forecast)

data(beersales)

plot(as.xts(beersales))
```


```{r}
bs <- as.xts(beersales)

train <- bs['1975/1989']
test <- bs['1990']
```


## 1A - Recursive

Use default options to resurvively forecast for 1990

```{r}
m <- auto.arima(train, seasonal = F)
plot(forecast(m, h=12))
```


```{r}
m <- auto.arima(train)
plot(forecast(m, h=12))
```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


