---
title: "31006: Assignment 3"
subtitle: "ARIMA Modeling"
author: "Scott Shepard"
date: "4/11/2019"
output: html_document
---

# Prompt

**ARIMA modeling:**

1. Use datasets from 1955 to 1968 to build an ARMA or ARIMA models for UN and GDP. Use auto.arima() from package forecast.  
2. Justify why you chose (ARMA or ARIMA) one over the other.  Note there will be 2 models, one for UN and another for GDP.  
3. Use the chosen UN and GDP models to forecast the UN and the GDP for 1969. Use forecast() from package forecast  
4. Compare your forecasts with the actual values using error = actual - estimate and plot the errors.  
5. Calculate the sum of squared error for each UN and GDP models.

**Regression - build regression models that use:**

1. UN as the independent variable and GDP as the dependent variable - use data from 1955 to 1968 to build the model. Forecast for 1969 and plot the errors as a percentage of the mean. Also calculate the sum of squared(error) as a percentage of the mean.  
2. GDP as the independent variable and UN as the dependent variable - use data from 1955 to 1968 to build the model.  Forecast for 1969 and plot the errors as a percentage of the mean. Also calculate the sum of squared(error) as a percentage of the mean of the actual values.  
3. Compare the 2 models using the sum of squared error as a percentage of the mean of the actual values - any reason to believe which should be the independent and the dependent variable ?  

## Setup & Data Loading

```{r}
library(openxlsx)

df <- read.xlsx("~/Datasets/31006/Unemployment_GDP_UK.xlsx", 1)

for( i in 2:nrow(df)) {
  if(is.na(df$Year[i])) {
    df$Year[i] <- df$Year[i-1]
  }
}

head(df)
```

# ARIMA Modeling

```{r, warning=FALSE}
library(zoo)
library(xts)

quarters <- as.yearqtr(paste(df$Year, df$Quarter, sep="-"))

UN  <- as.xts(df$UN,  order.by=quarters)
GDP <- as.xts(df$GDP, order.by=quarters)

par(mfrow=c(2,1))
plot.xts(UN)
plot.xts(GDP)
```

## 1. Build ARMA/ARIMA Models for UN & GDP for 1955-1968

```{r, message=FALSE}
library(forecast)

UN_arima <- auto.arima(UN['1955/1968'])
GDP_arima <- auto.arima(GDP['1955/1968'])

summary(UN_arima)
summary(GDP_arima)
```

## 2. Justify choice of ARMA or ARIMA

The I in ARIMA stands for integrated. It's used to difference the dataset 
first before applying an ARMA model. ARIMA models are useful when working on non-stationary time series. Both the GDP and UN time series are both 
non-stationary so using an ARIMA model instead of an ARMA model is 
appropriate.

```{r, message=FALSE, warning=FALSE}
library(tseries)

kpss.test(UN)
kpss.test(GDP)
```

The KPSS levels are both high enough to reject the null and accept the 
alternate that these series as non-stationary.

## 3. Forest UN & GDP for 1969

```{r, fig.width=12, fig.height=8}
UN_1969_forecast  <- forecast(UN_arima, h=4)
GDP_1969_forecast <- forecast(GDP_arima, h=4)

UN_1969_forecast
GDP_1969_forecast

par(mfrow=c(2,1))
plot(UN_1969_forecast)
plot(GDP_1969_forecast)
```

## 4. Compare forecasts with the actual values and plot errors

```{r}
UN_error  <- UN['1969']  - data.frame( UN_1969_forecast)$Point.Forecast
GDP_error <- GDP['1969'] - data.frame(GDP_1969_forecast)$Point.Forecast

par(mfrow=c(2,1))
plot(UN_error)
plot(GDP_error)
```

## 5. Calcuate sum of squared error for each model

```{r}
sum(UN_error^2)
sum(GDP_error^2)
```

# Regression

## 1. UN as ind. var and GDP as dep. var

```{r}
train <- df[df$Year < 1969,]
test <- df[df$Year == 1969,]

m1 <- lm(GDP ~ UN, train)

plot(test$GDP, predict(m1, newdata = test))
```

## 1. UN as dep. var and GDP as ind. var


```{r}
m2 <- lm(UN ~ GDP, train)

predict(m2, newdata = test)
```
