---
title: 'Week 1: Comparison of Correlations'
author: "Scott Shepard"
date: "4/1/2018"
output:
  pdf_document: default
  html_document: default
subtitle: MScA, Statistical Analysis (31007)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Comparison of correlations

Use the given simulated data to:
1. Read in the data set  
2. Plot the simulated pairs  
3. Calculate sample correlations and determination coefficients  
4. Interactive testing portion  

## Read the data

```{r read_data}
datapath <- "~/Dropbox/MScA/31007 - Stats Analysis/Assignment 1"
data <- read.csv(file.path(datapath, "Week1_Correlation_Comparison_Project_Data.csv"))
head(data)
```

## Plot the simulated pairs

```{r plot_pairs}
par(mfrow = c(2, 2))
with (data, { 
      plot(Case1.X, Case1.Y, ylim = c(-8, 8))
      plot(Case2.X, Case2.Y, ylim = c(-6, 6))
      plot(Case3.X, Case3.Y, ylim = c(-1, 3))
      plot(Case4.X, Case4.Y, ylim = c(0, 2.5))
})
```

## Calculate sample correlations and determination coefficients

```{r sample_cor_and_var}
with(data, {
  Correlation.Case1 <- cor(Case1.X, Case1.Y)
  Correlation.Case2 <- cor(Case2.X, Case2.Y)
  Correlation.Case3 <- cor(Case3.X, Case3.Y)
  Correlation.Case4 <- cor(Case4.X, Case4.Y)
  rbind(cbind(Correlation.Case1, 
              Correlation.Case2, 
              Correlation.Case3, 
              Correlation.Case4),
        cbind(Correlation.Case1^2,
              Correlation.Case2^2,
              Correlation.Case3^2,
              Correlation.Case4^2))
})
```

# Test. Relationship between slope and correlation

## Read the sample data
```{r test_read_data}
dat <- read.table(file.path(datapath, 'Week1_Test_Sample.csv'), header=TRUE)
```

## Calculate statistics
```{r test_calculate_stats}
sdX <- sd(dat$x)
sdY <- sd(dat$y)
cXY <- cor(dat$x, dat$y)
a   <- cXY * sdY / sdX
result <- data.frame(sdX=sdX, sdY=sdY, cXY=cXY, a=a)
result
```

## Write result to a file
```{r test_output}
write.table(result, file = file.path(datapath, 'result.csv'), row.names = F)
```

