---
title: "Course Assignment"
subtitle: "31007: Statistical Analysis"
author: "Scott Shepard"
date: "5/8/2018"
output: pdf_document
---

Completion:  
  - [x] Step 1  
  - [ ] Step 2  
  - [ ] Step 3  
  - [ ] Step 4  
  - [ ] Step 5  
  - [ ] Step 6  
  - [ ] Step 7  

```{r include=FALSE}
suppressWarnings(library(dplyr))
library(ggplot2)
```

## Step 1.

Read the data and visualize and get familiar with the variables.

```{r}
datapath <- "~/Dropbox/MScA/31007 - Stats Analysis/Course Project"

AssignmentData <- read.csv(
  file=file.path(datapath, "RegressionAssignmentData2014.csv"),
  row.names=1,
  header=TRUE,
  sep=",")

head(AssignmentData)
```

The first 7 variables (input variables) are the daily records of the US Treasury yields to maturity. The meaning of the variable Output will become clear later.

Plot the input variables.

```{r}
matplot(AssignmentData[,-c(8,9,10)],type='l')
```

Plot the input variables together with the output variable.

```{r}
matplot(AssignmentData[,-c(9,10)],type='l')
```

## Step 2

Estimate simple regression model with each of the input variables and the output variable given in AssignmentData.

### Three Month Bill

```{r}
m3M <- lm(Output1 ~ USGG3M, data=AssignmentData)
```

Check available names of fields returned by lm() and summary() functions

```{r names}
(names(m3M))
(names(summary(m3M)))
```

Analyze the summary

```{r summary_3_mo_model}
summary(m3M)
```

This variable, `USGG3M`, which trackes the rates of three month treasury bonds, 
is quite a good predictor of `Output1`. The p-value indicates a statistically 
significant slope. The adjusted R-squared is at 96% which is really high.

The key value of 2.5 for the estimated slope indicates that when all 

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m3M)$sigma^2)
```

There remains only a small portion of the total variance that is unexplained.
The ratio of these two numbers gives the unadjusted R^2 in summary.

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m3M$fitted.values,col="red")
```

### Six Month Bill

```{r}
m6M <- lm(Output1 ~ USGG6M, data=AssignmentData)

summary(m6M)
```

The six-month yield input does an even better job of predicting `Output1` than
the three-month yield did, explaining 97.4% of the variance.

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m6M)$sigma^2)
```

There is less unexplained variance with this input variable, which matches
the R^2 number in summary. The intercept is slightly different but the 
slope is about the same.

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m6M$fitted.values,col="red")
```

### Two Year Note

```{r}
m2Y <- lm(Output1 ~ USGG2YR, data=AssignmentData)

summary(m2Y)
```

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m2Y)$sigma^2)
```

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m2Y$fitted.values,col="red")
```

### Three Year Note

```{r}
m3Y <- lm(Output1 ~ USGG3YR, data=AssignmentData)

summary(m3Y)
```

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m3Y)$sigma^2)
```

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m3Y$fitted.values,col="red")
```

### Five Year Note

```{r}
m5Y <- lm(Output1 ~ USGG5YR, data=AssignmentData)

summary(m5Y)
```

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m5Y)$sigma^2)
```

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m5Y$fitted.values,col="red")
```

### Ten Year Note
 
```{r}
m10 <- lm(Output1 ~ USGG10YR, data=AssignmentData)

summary(m10)
```

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m10)$sigma^2)
```

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m10$fitted.values,col="red")
```

### Thirty Year Note

```{r}
m30 <- lm(Output1 ~ USGG30YR, data=AssignmentData)

summary(m30)
```

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m30)$sigma^2)
```

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m30$fitted.values,col="red")
```

### All Together Now

Collect all slopes and intercepts in one table and print this table. Try to do it in one line using apply() function.



```{r}
Output1 <- AssignmentData$Output1
t(apply(AssignmentData[,-c(8:10)], 2, function(x) lm(Output1~x)$coef))
```

# Step 3

Fit linear regression models using single output (column 8 Output1) as input and each of the original inputs as outputs.

Collect all slopes and intercepts in one table and print this table.
```{r}
t(apply(AssignmentData[,-c(8:10)], 2, function(y) lm(y~Output1)$coef))
```

# Step 4

Estimate logistic regression using all inputs and the data on FED tightening and easing cycles.

```{r}
AssignmentDataLogistic <- data.frame(data.matrix(AssignmentData, rownames.force="automatic"))
```

Prepare the easing-tightening data.
Make the easing column equal to 0 during the easing periods and NA otherwise.
Make the tightening column equal to 1 during the tightening periods and NA otherwise.

```{r}
# Create columns of easing periods (as 0s) and tightening periods (as 1s)
EasingPeriods<-AssignmentDataLogistic[,9]
EasingPeriods[AssignmentDataLogistic[,9]==1]<-0
TighteningPeriods<-AssignmentDataLogistic[,10]
# Check easing and tightening periods
cbind(EasingPeriods,TighteningPeriods)[c(550:560,900:910,970:980),]
```

Remove the periods of neither easing nor tightening.

```{r}
All.NAs<-is.na(EasingPeriods)&is.na(TighteningPeriods)
AssignmentDataLogistic.EasingTighteningOnly<-AssignmentDataLogistic
AssignmentDataLogistic.EasingTighteningOnly[,9]<-EasingPeriods
AssignmentDataLogistic.EasingTighteningOnly<-AssignmentDataLogistic.EasingTighteningOnly[!All.NAs,]
AssignmentDataLogistic.EasingTighteningOnly[is.na(AssignmentDataLogistic.EasingTighteningOnly[,10]),10]<-0
# Binary output for logistic regression is now in column 10
```

Plot the data and the binary output variable representing easing (0) and tightening (1) periods.

```{r}
matplot(AssignmentDataLogistic.EasingTighteningOnly[,-c(9,10)],type="l",ylab="Data and Binary Fed Mode")
lines(AssignmentDataLogistic.EasingTighteningOnly[,10]*20,col="red")
```

Estimate logistic regression with 3M yields as predictors for easing/tightening output.

```{r}
LogisticModel.TighteningEasing_3M<-glm(AssignmentDataLogistic.EasingTighteningOnly[,10]~
                        AssignmentDataLogistic.EasingTighteningOnly[,1],family=binomial(link=logit))
summary(LogisticModel.TighteningEasing_3M)
```

```{r}
matplot(AssignmentDataLogistic.EasingTighteningOnly[,-c(9,10)],type="l",ylab="Data and Fitted Values")
lines(AssignmentDataLogistic.EasingTighteningOnly[,10]*20,col="red")
lines(LogisticModel.TighteningEasing_3M$fitted.values*20,col="green")
```

```{r}
LogisticModel.TighteningEasing_All <-
  glm(Tightening ~ .,
      data=AssignmentDataLogistic.EasingTighteningOnly[,c(1:7,10)],
      family=binomial(link=logit))
```

```{r}
summary(LogisticModel.TighteningEasing_All)$aic
```

```{r}
summary(LogisticModel.TighteningEasing_All)$coefficients[,c(1,4)]
```

```{r}
matplot(AssignmentDataLogistic.EasingTighteningOnly[,-c(9,10)],type="l",ylab="Results of Logistic Regression")
lines(AssignmentDataLogistic.EasingTighteningOnly[,10]*20,col="red")
lines(LogisticModel.TighteningEasing_All$fitted.values*20,col="green")
```

**Interpret the coefficients of the model and the fitted values.**

### TODO

Calculate and plot log-odds and probabilities. Compare probabilities with fitted values.

```{r}
# Calculate odds
Log.Odds<-predict(LogisticModel.TighteningEasing_All)
plot(Log.Odds,type="l")
```

```{r}
Probabilities<-1/(exp(-Log.Odds)+1)
plot(LogisticModel.TighteningEasing_All$fitted.values,type="l",ylab="Fitted Values & Log-Odds")
lines(Probabilities,col="red")
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

