---
title: "Course Assignment"
subtitle: "31007: Statistical Analysis"
author: "Scott Shepard"
date: "5/8/2018"
output: pdf_document
---

Completion:  
  - [x] Step 1  
  - [ ] Step 2  
  - [ ] Step 3  
  - [ ] Step 4  
  - [ ] Step 5  
  - [ ] Step 6  
  - [ ] Step 7  

```{r include=FALSE}
suppressWarnings(library(dplyr))
library(ggplot2)
```

## Step 1.

Read the data and visualize and get familiar with the variables.

```{r}
datapath <- "~/Dropbox/MScA/31007 - Stats Analysis/Course Project"

AssignmentData <- read.csv(
  file=file.path(datapath, "RegressionAssignmentData2014.csv"),
  row.names=1,
  header=TRUE,
  sep=",")

head(AssignmentData)
```

The first 7 variables (input variables) are the daily records of the US Treasury yields to maturity. The meaning of the variable Output will become clear later.

Plot the input variables.

```{r}
matplot(AssignmentData[,-c(8,9,10)],type='l')
```

Plot the input variables together with the output variable.

```{r}
matplot(AssignmentData[,-c(9,10)],type='l')
```

## Step 2

Estimate simple regression model with each of the input variables and the output variable given in AssignmentData.

### Three Month Bill

```{r}
m3M <- lm(Output1 ~ USGG3M, data=AssignmentData)
```

Check available names of fields returned by lm() and summary() functions

```{r names}
(names(m3M))
(names(summary(m3M)))
```

Analyze the summary

```{r summary_3_mo_model}
summary(m3M)
```

This variable, `USGG3M`, which trackes the rates of three month treasury bonds, 
is quite a good predictor of `Output1`. The p-value indicates a statistically 
significant slope. The adjusted R-squared is at 96% which is really high.

The key value of 2.5 for the estimated slope indicates that when all 

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m3M)$sigma^2)
```

There remains only a small portion of the total variance that is unexplained.
The ratio of these two numbers gives the unadjusted R^2 in summary.

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m3M$fitted.values,col="red")
```

### Six Month Bill

```{r}
m6M <- lm(Output1 ~ USGG6M, data=AssignmentData)

summary(m6M)
```

The six-month yield input does an even better job of predicting `Output1` than
the three-month yield did, explaining 97.4% of the variance.

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m6M)$sigma^2)
```

There is less unexplained variance with this input variable, which matches
the R^2 number in summary. The intercept is slightly different but the 
slope is about the same.

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m6M$fitted.values,col="red")
```

### Two Year Note

```{r}
m2Y <- lm(Output1 ~ USGG2YR, data=AssignmentData)

summary(m2Y)
```

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m2Y)$sigma^2)
```

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m2Y$fitted.values,col="red")
```

### Three Year Note

```{r}
m3Y <- lm(Output1 ~ USGG3YR, data=AssignmentData)

summary(m3Y)
```

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m3Y)$sigma^2)
```

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m3Y$fitted.values,col="red")
```

### Five Year Note

```{r}
m5Y <- lm(Output1 ~ USGG5YR, data=AssignmentData)

summary(m5Y)
```

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m5Y)$sigma^2)
```

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m5Y$fitted.values,col="red")
```

### Ten Year Note
 
```{r}
m10 <- lm(Output1 ~ USGG10YR, data=AssignmentData)

summary(m10)
```

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m10)$sigma^2)
```

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m10$fitted.values,col="red")
```

### Thirty Year Note

```{r}
m30 <- lm(Output1 ~ USGG30YR, data=AssignmentData)

summary(m30)
```

```{r}
c(Total.Variance=var(AssignmentData[,8]),Unexplained.Variance=summary(m30)$sigma^2)
```

```{r}
matplot(AssignmentData[,8],type="l",xaxt="n")
lines(m30$fitted.values,col="red")
```

### All Together Now

Collect all slopes and intercepts in one table and print this table. Try to do it in one line using apply() function.

```{r}
models <- list(m3M, m6M, m2Y, m3Y, m5Y, m10, m30)
models_table <- plyr::ldply(models, function(m) {
  coefficients <- coef(m)
  var_name <- names(coefficients)[2]
  data.frame(Input=var_name,Intercept=coefficients[1], Slope=coefficients[1])
})
row.names(models_table) <- models_table$Input
models_table <- models_table[,-1]
models_table
```

# Step 3

Fit linear regression models using single output (column 8 Output1) as input and each of the original inputs as outputs.

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

