---
title: "Picking the Best Model for Genre Prediction"
author: "Scott Shepard"
subtitle: "Trees, Random Forest, and Multinomial Logit"
date: "8/12/2018"
output: pdf_document
---

# Introduction

In this project, we used the Free Music Archive dataset to first
predict missing top-level genres from the echonest dataset and
then used those predictions to enhance a playlist generator
algorithm. We used the echonost audio features for both the
parts of the project.

Our final model was a simple classificaiton tree with a rather
deep structure. We compared this model against other algorthims
we learned about in Data Mining, specifically the Random Forest,
Multinomial Logit, and Linear Discriminant Anaysis. This 
document examains other models we could have used.

```{r libraries}
suppressMessages(library(dplyr))
library(ggplot2)
suppressMessages(library(gridExtra))
library(stringr)
library(reshape2)
library(rpart)
suppressMessages(library(MASS))
library(nnet)
suppressMessages(library(randomForest))
```

The dataset is a cleaned version of what can be found on
[FMA github page](https://github.com/mdeff/fma).

```{r clean_data}
path <- "~/Dropbox/MScA/31008 - Data Mining/Project/"
df <- read.csv(file.path(path, "Project_Data.csv"))

names(df)[names(df) == "genre_top"] <- "genre"
df$genre <- as.character(df$genre)
df$genre <- factor(df$genre)

df$tempo_scaled <- scales::rescale(df$tempo)
```

```{r}
df$genre <- as.character(df$genre)
df <- df[! df$genre %in% c("", "International", "Experimental"), ]
df[df$genre %in% c("Folk", "Blues"), ]$genre <- c("Folk/Blues")
df$genre <- factor(df$genre)
  
table(df$genre)
```

```{r}
ggplot(df, aes(x=danceability)) + 
  geom_histogram(fill="#1db954", color="black") + 
  ggthemes::theme_fivethirtyeight() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title="Danceability!")

ggplot(df, aes(x=acousticness)) + 
  geom_histogram(fill="#1db954", color="black") + 
  ggthemes::theme_fivethirtyeight() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(title="Acousticness")
```

# Model Training

Split the data

```{r}
splitData <- function(data, smp_size) {
  # Take a dataset and split it randomly into train and test sets
  # The size of the training set is the smp_size argument
  train_ind <- sample(seq_len(nrow(data)), size = smp_size)
  train <- data[ train_ind, ]
  test  <- data[-train_ind, ]
  
  list("train"=train, "test"=test)
}
```

```{r data_holdout}
set.seed(1)
l <- splitData(df, round(nrow(df) * 0.7))
train <- l$train
holdout <- l$test
```

## Simple Classification Tree

The tree model parameters come from Anna Whelan.

```{r}
Simple.Tree = rpart(genre ~ acousticness + danceability + 
                      energy + instrumentalness + liveness +
                      speechiness + tempo + valence,
                    data=train, control=rpart.control(cp=9.6360e-04, minsplit=25, xval=20, maxsurrogate=0))

train$Genre.Tree <- predict(Simple.Tree, type="class")
```

The main metric I used is accuracy. The question accuracy 
answers is, "Given a prediction, what is the liklihood that
prediction is correct?" We want high, across the board accuracy 
measures in holdout.

```{r}
accuracyTable <- function(df, pred.col) {
  dots <- list(
    lazyeval::interp(~ mean(var == genre), var=as.name(pred.col)),
    ~n())
  group_by_(df, pred.col) %>%
    summarize_(.dots=setNames(dots, c("pct", "n")))
}

correctClassifyTable <- function(df, pred.col) {
  dots <- list(
    lazyeval::interp(~ mean(var == genre), var=as.name(pred.col)),
    ~n())
  group_by(df, genre) %>%
    summarize_(.dots=setNames(dots, c("pct", "n")))
}

accuracyPlot <- function(df, pred.col) {
  accuracy <- accuracyTable(df, pred.col=pred.col)
  
  ggplot(accuracy, aes_string(x=pred.col, y="pct")) + 
    geom_bar(stat="identity", fill="blue4") + 
    geom_text(aes(y=pct, label=paste0(round(pct*100,1), "%")),
              vjust=1.5, color="white") +
    geom_text(aes(y=0, label=paste0("N=",n)),
              vjust=-0.5, color="white") +
    scale_y_continuous(labels=scales::percent, 
                       limits = c(0,1)) +
    ggthemes::theme_fivethirtyeight() +
    theme(axis.title.x = element_blank()) +
    labs(title="Preciction Accuracy",
         subtitle="Given a Prection, How Likely is it to be true?",
         y="Percent of Values Predicted Correctly")
}

correctClassifyPlot <- function(df, pred.col) {
  correct_classify <- correctClassifyTable(df, pred.col=pred.col)
  
  ggplot(correct_classify, aes(x=genre, y=pct)) + 
    geom_bar(stat="identity", fill="blue4") + 
    geom_text(aes(y=pct, label=paste0(round(pct*100,1), "%")),
              vjust=1.5, color="white") +
    scale_y_continuous(labels=scales::percent,limits = c(0,1)) +
    ggthemes::theme_fivethirtyeight() +
    theme(axis.title.x = element_blank()) +
    labs(title="Correct Classification",
         subtitle="Given a track, how likely is it to be correctly classified?",
         y="Percent of Observations Correctly Classified")
}
```

```{r}
holdout$Genre.Tree <- predict(Simple.Tree, type="class", 
                              newdata = holdout)

tree_plot = accuracyPlot(holdout, "Genre.Tree") + 
  ggtitle("Accuracy of Classification Tree")
tree_plot
```

## Multinomial Logit

```{r}
Multi.Logit <- nnet::multinom(
  genre ~ acousticness + danceability + energy + 
    instrumentalness + liveness + speechiness + tempo + valence,
  data=train)

holdout$Genre.Multilogit <- predict(Multi.Logit, newdata=holdout)

logit_plot = accuracyPlot(holdout, "Genre.Multilogit") + 
  ggtitle("Accuracy of Multi Logistic")
logit_plot
```

## LDA & QDA

```{r}
LDA.Model <- MASS::lda(
  genre ~ ., train[,c('genre','acousticness','danceability',
                      'energy','instrumentalness','liveness',
                      'speechiness','tempo','valence')])
holdout$Genre.LDA <- predict(LDA.Model, newdata=holdout)$class
lda_plot = accuracyPlot(holdout, "Genre.LDA") + 
  ggtitle("Accuracy of LDA")
lda_plot
```

```{r}
QDA.Model <- MASS::qda(
  genre ~ . ,train[,c('genre','acousticness','danceability'
                      ,'energy','instrumentalness','liveness',
                      'speechiness','tempo','valence')])
holdout$Genre.QDA <- predict(QDA.Model, newdata=holdout)$class
accuracyPlot(holdout, "Genre.QDA") + ggtitle("Accuracy of QDA")
```

## Random Forest

```{r}
Model.RF <- randomForest::randomForest(
  genre ~ acousticness + danceability + energy + instrumentalness +
    liveness + speechiness + tempo + valence, 
  data=train)

holdout$Genre.RF <- predict(Model.RF, newdata = holdout)

rf_plot = accuracyPlot(holdout, "Genre.RF") + 
  ggtitle("Accuracy of Random Forest")
rf_plot
```

# All plots together
